\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,booktabs,geometry,graphicx}
\geometry{margin=1in}

\title{Does the Generated-Regressor Correction Improve the Plug-in Estimator?}
\date{}

\begin{document}
\maketitle

\section{Testing procedure}

We test two questions empirically via Monte Carlo simulation:
\begin{enumerate}
\item Does the correction help when the selection-induced correlation is varied (by varying $\gamma$)?
\item Does the correction help at small sample sizes?
\end{enumerate}

\paragraph{Data generating process.}
\begin{align*}
x &\sim N(0,1), \quad a \sim \pm\,\mathrm{Exp}(1), \quad \eta \mid a \sim N(a,1), \\
w \mid \eta &\sim N\!\left(\tfrac{2e^\eta}{1+e^\eta},\; 0.1\right), \quad
\nu \mid a \sim N\!\left(\tfrac{3e^a}{1+e^a},\; 1\right), \\
q &= \gamma\,x + \eta, \quad s = \mathbf{1}\{q > 0\}, \quad y = s\,w + \beta\,x + \nu.
\end{align*}
Here $\beta = 2$ throughout. The nonparametric component is approximated by a cubic B-spline basis with $k_n = \max(4,\,\lfloor n^{1/3} \rceil)$ knots on fixed support $[-2.8,\,2.8]$, with out-of-support values clipped to the boundary.

\paragraph{Experiment 1: Varying $\gamma$.}
Fix $n = 50{,}000$ and $R = 500$ replications. Sweep $\gamma \in \{0.3,\, 0.7,\, 1.0,\, 1.5,\, 2.0,\, 3.0,\, 5.0,\, 10.0\}$. For each $\gamma$, compare the Robinson and naive estimators using the same data in each replication. To test whether amplifying the first-stage error $\delta$ changes the picture, we also run with the first stage estimated from only $n/10$ observations (amplifying $\delta$ by a factor of $\approx\!\sqrt{10} \approx 3.2$).

Report: empirical $\mathrm{Cor}(x, \hat\eta \mid \text{treated})$, bias, standard deviation, and variance ratio $\mathrm{Var}(\hat\alpha_{\mathrm{Rob}}) / \mathrm{Var}(\hat\alpha_{\mathrm{Naive}})$.

\paragraph{Experiment 2: Varying $n$.}
Fix $\gamma = 2$ and $R = 500$. Sweep $n \in \{500,\, 1000,\, 2000,\, 5000,\, 10{,}000,\, 50{,}000,\, 100{,}000\}$. Report variance ratio and $\sqrt{n} \cdot \mathrm{Std}$ for each method.

\section{Results}

\subsection{Experiment 1: Varying $\gamma$ (selection-induced correlation)}

\begin{table}[h]
\centering
\caption{Robinson vs.\ Naive at $n = 50{,}000$, full first-stage ($n_{\mathrm{fs}} = n$).}
\label{tab:gamma_full}
\begin{tabular}{r r r r r r r}
\toprule
$\gamma$ & $\mathrm{Cor}(x,\hat\eta|\mathrm{Tr})$ & Bias$_R$ & Bias$_N$ & Std$_R$ & Std$_N$ & VarRatio \\
\midrule
0.3  & $-0.17$ & 0.187 & 0.386 & 0.527 & 0.406 & 1.69 \\
0.7  & $-0.34$ & 0.233 & 0.191 & 0.468 & 0.421 & 1.24 \\
1.0  & $-0.41$ & 0.007 & 0.018 & 0.228 & 0.194 & 1.38 \\
1.5  & $-0.45$ & 0.011 & 0.011 & 0.028 & 0.028 & 1.00 \\
2.0  & $-0.45$ & 0.013 & 0.013 & 0.021 & 0.021 & 1.00 \\
3.0  & $-0.40$ & 0.013 & 0.013 & 0.018 & 0.018 & 1.00 \\
5.0  & $-0.30$ & 0.010 & 0.010 & 0.016 & 0.016 & 1.00 \\
10.0 & $-0.17$ & 0.006 & 0.006 & 0.016 & 0.016 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Robinson vs.\ Naive at $n = 50{,}000$, amplified first-stage error ($n_{\mathrm{fs}} = n/10$).}
\label{tab:gamma_amplified}
\begin{tabular}{r r r r r r r}
\toprule
$\gamma$ & $\mathrm{Cor}(x,\hat\eta|\mathrm{Tr})$ & Bias$_R$ & Bias$_N$ & Std$_R$ & Std$_N$ & VarRatio \\
\midrule
0.3  & $-0.18$ & 0.233 & 0.376 & 0.535 & 0.441 & 1.47 \\
0.7  & $-0.34$ & 0.250 & 0.208 & 0.472 & 0.417 & 1.28 \\
1.0  & $-0.41$ & 0.037 & 0.037 & 0.207 & 0.185 & 1.26 \\
1.5  & $-0.45$ & 0.012 & 0.012 & 0.028 & 0.028 & 1.02 \\
2.0  & $-0.45$ & 0.013 & 0.013 & 0.021 & 0.021 & 1.00 \\
3.0  & $-0.40$ & 0.012 & 0.012 & 0.018 & 0.018 & 1.00 \\
5.0  & $-0.31$ & 0.011 & 0.011 & 0.017 & 0.017 & 1.00 \\
10.0 & $-0.17$ & 0.007 & 0.007 & 0.017 & 0.017 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

For $\gamma \geq 1.5$, the variance ratio is indistinguishable from 1.00, even when the first-stage error is amplified by a factor of $\sim\!3$. The correction has no detectable effect on the plug-in estimator.

For small $\gamma$ ($\leq 1.0$), both methods exhibit large bias and variance (Std $\sim 0.2$--$0.5$ versus $\sim 0.02$ at $\gamma = 2$), and the correction actually \emph{increases} variance (VarRatio $> 1$). This regime is not a failure of the correction per se; rather, the plug-in estimator itself breaks down because small $\gamma$ implies poor overlap between the treated and control $\eta$ distributions, making the cross-group evaluation $\hat h_{\mathrm{Tr}}(\eta) - \hat h_{\mathrm{Con}}(\eta)$ unreliable.

Amplifying $\delta$ does not change the qualitative picture. The variance ratios at $\gamma \geq 1.5$ remain near 1.00, confirming that the generated-regressor error is simply not a binding constraint for the plug-in.

\subsection{Experiment 2: Varying $n$ at $\gamma = 2$}

\begin{table}[h]
\centering
\caption{Robinson vs.\ Naive across sample sizes, $\gamma = 2$.}
\label{tab:n_sweep}
\begin{tabular}{r r r r r r r r}
\toprule
$n$ & $k_n$ & Bias$_R$ & Bias$_N$ & Std$_R$ & Std$_N$ & VarRatio & $\sqrt{n}\!\cdot\!\mathrm{Std}_N$ \\
\midrule
500     &  8 & 0.018 & 0.018 & 0.278 & 0.247 & 1.26 & 5.53 \\
1{,}000   & 10 & 0.013 & 0.015 & 0.190 & 0.176 & 1.16 & 5.56 \\
2{,}000   & 13 & 0.015 & 0.018 & 0.114 & 0.111 & 1.06 & 4.98 \\
5{,}000   & 17 & 0.011 & 0.012 & 0.067 & 0.067 & 1.01 & 4.74 \\
10{,}000  & 22 & 0.014 & 0.014 & 0.045 & 0.045 & 1.00 & 4.49 \\
50{,}000  & 37 & 0.011 & 0.011 & 0.021 & 0.021 & 1.01 & 4.70 \\
100{,}000 & 46 & 0.012 & 0.012 & 0.015 & 0.015 & 1.00 & 4.57 \\
\bottomrule
\end{tabular}
\end{table}

At small $n$, the correction is actively harmful: at $n = 500$, Robinson has 26\% higher variance than the naive estimator. The penalty decreases monotonically with $n$ and is negligible by $n = 5{,}000$. The mechanism is multicollinearity: the correction adds a regressor (the shifted-basis difference) that is nearly collinear with $x$, inflating the variance of all OLS coefficients including $\hat\omega$. At small $n$ with few basis functions, this collinearity is most damaging.

\section{Conclusion}

The generated-regressor correction does not reduce the asymptotic variance of the plug-in estimator $\hat\alpha$ in this DGP. The theoretical reason is that the plug-in depends on the spline coefficients $\hat\omega$, not on $\hat\beta$, and the bias from the generated regressor is absorbed almost entirely by $\hat\beta$ due to the approximate orthogonality of $x$ and $\Phi(\hat\eta)$. The selection-induced within-subsample correlation between $x$ and $\hat\eta$ is the only channel through which the correction could help, but its effect is empirically undetectable---even when $\delta$ is artificially amplified. At small sample sizes, the correction is counterproductive due to the multicollinearity it introduces.

\end{document}
